### 2022-04-25 outputs/paper.Rmd
simulation

---
title: "Evolution of income (equality) in Canada"
author: 
  - Alicia Yang
thanks: "Code and data are available at: https://github.com/Alicia-y/Telling-stories-with-data-final-paper"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "People have claimed that individuals who are experiencing mental illness are more likely to be involved in violent incidents and crime activities either as victims or as criminals. To investigate more on this issue, this paper aims to explore how the mental health issue interacts with the reported crimes in Canada. By utilizing the data from Statistics Canada, we found that increase in mental health cases does have a positive correlation with the crime rates in Canada. More specifically, teenagers suffering from mental illness are much more likely to be involved in criminal activities. In addition, implications behind this correlation is discussed and possible solutions is proposed."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set up libraries
library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

```

\newpage

# Introduction

In recent decades the world as a whole is developing fast economically, and one evidence of which is the rise in income level. As more countries become more developed in terms of their economy and eventually overcome their issues of poverty, their governments have gradually shift their goal to dealing with the problem of economic inequality. 
It concerns how the wealth and income of the population is distributed. Although we are generally living higher quality lives and getting higher incomes, another problem is being put forward by people who are claiming that the gap between the rich and poor is getting larger due to technological advancement, globalization, countries' governmental structure and policies, which betters off the people who are already wealthy by making it easy for them to seek for more opportunities, thus, snowballing their assets faster. Income disparity can be a huge issue. It makes the income per capita of a country less reflective of its people's true living quality, because income per capita does not tell the distribution of the income. The richest people may pull up the statistics about the average/median income level, but it doesn't mean the country is performing well economically as a whole. In addition to that, income inequality also have other potential negative effects. It could reduce the stability of the economy and increase the risk of financial crisis. It could also result in corruption, misallocation of resources as the rich is empowered economically, socially and even politically.
Canada is one of the country in the world that provides the most benefits and welfare, yet people still claimed that the economic inequality is rising along with the income over the past 2 decades(2). Therefore, I'm interested in whether this is a serious issue in Canada. If so, in which regions and what period of time is it a serious issue? And is it possible to predict how the disparity of income change in future?

The data used in this report is obtained from Open Government Data of Statistics Canada. The dataset is called "Upper income limit, income share and average income by economic family type and income decile", and it provides different kinds information about the income level of the households of Canada between 1976 and 2020 by the Canadian Income Survey. In this dataset, I'm interested in the year the data is from, the geographical location, the income decile so that we could observe the difference between the lowest and highest decile, and the actual value of the income. In section \@ref(data), a cleaned dataset is obtained

Associated with

Difficult and challenging to deal with

 In section \@ref(data), the two datasets are discussed and cleaned to perform more accurate analysis. Visualizations in the forms of tables and graphs are presented to help to explain the possible correlation between mental illness cases and reported crime cases in Canada each year. In section \@ref(model), a model is constructed to justify the relationship and to make future predictions. All findings related to this topic is presented in section \@ref(results). A discussion is carried out in section \@ref(discussion) on the implications of the findings, possible solutions to this issue, as well as the limitations of this paper.

One of the reason why mental health is such a serious issue is that it's hard to identify and treat. Most patients undergo long-term treatments before fully recovered. To provide possible resolutions, measures are taken in recent years for the patients of mental illness, such as opening different types of community support services to help them cope with this issue. I will also look into the effectiveness of these support services and programs and see if they actually helped the ones suffering from mental health issues.

# Data

## Dataset Description and Methodology

The dataset, which is a summary of the income-related statistics in Canada between 1976 and 2020, is obtained from the Open Government Portal of Statistics Canada. The source of the data is from the Canadian Income Survey conducted annually. The survey covers all the possible population in Canada except for the people and households living in remote areas or indigenous settlements. It doesn't really affect the outcome of the survey, because these people only take up less than 2 percent of the population(). The selected respondents were drawn from the Labour Force Survey samples, which is based on a stratified probability sampling(). To reduce some non-sampling errors, telephone interviews ware conducted prior to the main survey to increase the response rate of the selected participants Once they give consent to completing the survey, they will be able to complete it in an online form. Respondents are protected from the confidentiality rules, so that their privacy won't be disclosed after the survey is made public.
One possible bias came with the survey methodology was the over-coverage of the units that are not the target population and the under-coverage of certain sub-population that should be included in the survey, such as certain remote areas.

The original dataset contains 173043 observations and 18 variables with all kinds of information about income level statistics. This report wants to focus on exploring the trend of income inequality and investigating the possible factors that may affect income inequality. Thus, I will be interested in the variables "Year", "Geographical Location", "Income decile", "Income" which tells how the income level of different regions of Canada has evolved throughout these years. R [@citeR], and R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "knitr" [@citeknitr], "dplyr" [@citedplyr], and "kableExtra" [@citekableExtra] are utilized to create an extract of the cleaned dataset (Table \@ref(tab:dataextract)).

```{r, echo=FALSE, message=FALSE}
# Load the cleaned data
income <- read_csv(here::here("inputs/cleaned_data.csv"))
```

```{r dataextract, echo=FALSE, message=FALSE}
income %>%
  slice(1:10) %>%
  kable(
    caption = "Extracting the first ten rows from the Income data",
col.names=c("Year", "Geographical location", "Income decile", "Income", "Income range"),
linesep = "", digits = 1, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)
```

Table \@ref(tab:dataextract) shows the first ten rows of the cleaned dataset. It contains 6395 variables and 5 variables in total. As we known, the information from the dataset is extracted from the Canada Income Survey. Variable "Year" indicates the year of the survey. Variable "Geographical location" indicates where the group of respondents are from. Variable "Income decile" gives more information about the variable "Income", by telling us

## Data Visualization

```{r, echo=FALSE, message=FALSE}

income %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical Location` == "Canada") %>%
  ggplot(aes(x=Year, y=Income))+geom_point()+theme_minimal()+ylab("Average after-tax income in Canada")
```

```{r, echo=FALSE, message=FALSE}
income %>%
  filter(
         `Income decile` == "Total deciles") %>%
  ggplot(aes(x=Year, y=`Income range`, colour = `Geographical Location`))+geom_point()+theme_minimal()
```

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

https://www.nytimes.com/2020/07/01/opinion/economic-inequality-moral-philosophy.html#:~:text=Enough%20economic%20inequality%20can%20transform,society%20ruled%20by%20the%20rich.&text=Large%20inequalities%20of%20inherited%20wealth,and%20undercuts%20equality%20of%20opportunity.

(2)https://www.conferenceboard.ca/hcp/hot-topics/canInequality.aspx#:~:text=Income%20inequality%20in%20Canada%20has,income%20Canadians%20also%20lost%20share.

https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5200